Forecasting Influenza Hospitalizations Using a Bayesian Hierarchical Nonlinear Model with Discrepancy

### 2025-06-30 manuscript_res.tex
fix full fit

\documentclass{article}

\usepackage{PRIMEarxiv}

\usepackage[utf8]{inputenc} % allow utf-8 input
% \usepackage[T1]{fontenc}    % use 8-bit T1 fonts
\usepackage[linktocpage=true]{hyperref}       % hyperlinks
\usepackage{url}            % simple URL typesetting
\usepackage{booktabs}       % professional-quality tables
\usepackage{amsfonts}       % blackboard math symbols
\usepackage{nicefrac}       % compact symbols for 1/2, etc.
\usepackage{microtype}      % microtypography
% \usepackage{lipsum}
\usepackage{fancyhdr}       % header
\usepackage{graphicx}       % graphics
\graphicspath{{media/}}     % organize your images and other figures under media/ folder

% Packages added by Spencer
\usepackage{amsmath,amsthm,amssymb,scrextend}
\usepackage{appendix}
\AtBeginEnvironment{appendices}{\renewcommand\thesection{\Alph{section}}}
\usepackage{enumerate}
\usepackage{csquotes}
% \usepackage{subcaption}
\usepackage{amssymb}
% \usepackage{bbm}
\usepackage[bb=pazo]{mathalpha}
\usepackage{bm}
\usepackage{newtxmath} %for making indicator function
\usepackage{booktabs} %for the table in the analysis section
% \usepackage{enumerate} %for the short algorithm
\usepackage{subfigure} %for making plots with multiple images
\usepackage{tabularx} %for the spacing in tabular

\pagestyle{empty}

\newcommand\Bbbbone{%
  \ifdefined\mathbbb%
    \mathbbb{1}%
  \else%
    \boldsymbol{\mathbb{1}}%
  \fi}

% \usepackage{xr}
% \externaldocument{supplementary}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% \usepackage[pdftex]{graphicx}
% Standard, old-style thesis
% \usepackage{isutraditional}   
% \chaptertitle
% Old-style, thesis numbering down to subsubsection
% \alternate
% \usepackage{rotating}
% Bibliography without numbers or labels
\usepackage{natbib}
% Use the following line if you want square brackets and numbering system
%\usepackage[square, numbers]{natbib}

% this package is for getting certain math scripts like the 1 needed for an indicator function
% \usepackage{bbm}

% using this package to make a better looking list
\usepackage{enumerate}
% try this one for a list as well. may be better
% \usepackage[font=itshape]{quoting}

% package for tables

% \renewcommand{\bibfont}{\setstretch{1}\selectfont}
% \setlength{\bibsep}{13.2pt}

\usepackage{chapterbib}
%\renewcommand\bibsection{\section*{\ref name}}
%\renewcommand\bibsection{\section*}
\renewcommand{\bibsection}{\section{References}}

%\bibliographystyle{apa}  % changed from apa, abbrv, unsrtnat
%\includeonly{titletoc,chapter1}
%Optional Package to add PDF bookmarks and hypertext links
% \usepackage[pdftex,hypertexnames=false,linktocpage=true]{hyperref}
\hypersetup{colorlinks=true,linkcolor=blue,anchorcolor=blue,citecolor=blue,filecolor=blue,urlcolor=blue,bookmarksnumbered=true,pdfview=FitB}

%\usepackage{titletoc}
% \usepackage{hyperref}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% Packages/commands added by Jarad
\usepackage{xcolor}
\newcommand{\jarad}[1]{{\color{red} Jarad: #1}}
\newcommand{\spencer}[1]{{\color{blue} Spencer: #1}}

%Header
\pagestyle{fancy}
\thispagestyle{empty}
\rhead{ \textit{ }}

% Update your Headers here
\fancyhead[LO]{Bayesian nonlinear flu hospitalizations forecasting}
% \fancyhead[RE]{Firstauthor and Secondauthor} % Firstauthor et al. if more than 2 - must use \documentclass[twoside]{article}

%% Title
\title{Forecasting Influenza Hospitalizations Using a Bayesian Hierarchical Nonlinear Model with Discrepancy
%%%% Cite as
%%%% Update your official citation here when published 
% \thanks{\textit{\underline{Citation}}: 
% \textbf{Authors. Title. Pages.... DOI:000000/11111.}} 
}

\author{
  Spencer Wadsworth \\
  Iowa State University \\
  Ames, Iowa\\
  \texttt{sgw96@iastate.edu} \\
  %% examples of more authors
   \And
  Jarad Niemi \\
  Iowa State University \\
  Ames, Iowa\\
  \texttt{niemi@iastate.edu} \\
  %% \AND
  %% Coauthor \\
  %% Affiliation \\
  %% Address \\
  %% \texttt{email} \\
  %% \And
  %% Coauthor \\
  %% Affiliation \\
  %% Address \\
  %% \texttt{email} \\
  %% \And
  %% Coauthor \\
  %% Affiliation \\
  %% Address \\
  %% \texttt{email} \\
}

\begin{document}
\maketitle
\begin{abstract}
The annual influenza outbreak leads to significant public health and economic burdens making it desirable to have prompt and accurate probabilistic forecasts of the disease spread. The United States Centers for Disease Control and Prevention (CDC) hosts annually a national flu forecasting competition which has led to the development of a variety of flu forecast modeling methods. For the first several years of the competition, the target to be forecast was weekly percentage of patients with an influenza-like illness (ILI), but in 2021 the target was changed to weekly hospitalization counts. Reliable state and national hospitalization data has only been available since 2021, but for ILI the data has been available since 2010 and has been successfully forecast for several seasons.
In this manuscript, we introduce a two component modeling framework for forecasting weekly hospitalizations utilizing both hospitalization data and ILI data. The first component is for modeling ILI data using a dynamic nonlinear Bayesian hierarchical model. The second component is for modeling hospitalizations as a function of ILI. For hospitalization forecasts, ILI is first forecasted and then hospitalizations are forecast with ILI forecasts used as a linear or quadratic predictor. In a simulation study, two ILI forecast models, including one similar to the winning model for two seasons of the CDC forecast competition from Osthus et al.
% \cite[]{osthus2019dynamic} 
% 
and a nonlinear Bayesian hierarchical model from Ulloa 
% \cite[]{ulloa2019}
are compared. Also assessed is the usefulness of including a systematic model discrepancy term in the ILI model. Forecasts of state and national hospitalizations for the 2023-24 flu season are made, and different modeling decisions are compared. We found that including a discrepancy component in the ILI model tends to improve forecasts during certain weeks of the year. We also found that other modeling decisions such as the exact nonlinear function to be used in the ILI model or the error distribution for hospitalization models may or may not be better than other decisions, depending on the season, location, or week of the forecast.
\end{abstract}

% \begin{abstract}
% The annual influenza outbreak leads to significant public health and economic burdens making it desirable to have prompt and accurate probabilistic forecasts of the disease spread. The United States Centers for Disease Control and Prevention (CDC) hosts annually a national flu forecasting competition which has led to the development of a variety of flu forecast modeling methods. Beginning in 2013, the target to be forecast was weekly percentage of patients with an influenza-like illness (ILI), but in 2021 the target was changed to weekly hospitalizations. Reliable hospitalization data has only been available since 2021, but ILI data has been available since 2010 and has been successfully forecast for several seasons. In this manuscript, we introduce a two component modeling framework for forecasting hospitalizations utilizing both hospitalization and ILI data. The first component is for modeling ILI data using a nonlinear Bayesian model. The second component is for modeling hospitalizations as a function of ILI. For hospitalization forecasts, ILI is first forecast then hospitalizations are forecast with ILI forecasts used as a predictor. In a simulation study, the hospitalization forecast model is assessed and two previously successful ILI forecast models are compared. Also assessed is the usefulness of including a systematic model discrepancy term in the ILI model. Forecasts of state and national hospitalizations for the 2023-24 flu season are made, and different modeling decisions are compared. We found that including a discrepancy component in the ILI model tends to improve forecasts during certain weeks of the year. We also found that other modeling decisions such as the exact nonlinear function to be used in the ILI model or the error distribution for hospitalization models may or may not be better than other decisions, depending on the season, location, or week of the forecast.
% \end{abstract}

% \begin{keyword}[class=MSC]
% \kwd[Primary ]{00X00}
% \kwd{00X00}
% \kwd[; secondary ]{00X00}
% \end{keyword}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% `\ ' is used here because TeX ignores    %%
%% spaces after text commands.              %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}

@@

The administration of FluSight saw few changes during the first seven seasons, but the onset of the COVID-19 pandemic and subsequent developments for COVID-19 forecasting led to major modifications. As a result of the COVID-19 pandemic which began during the 2019 flu season, the typical flu outbreak behavior was altered between the 2019 and 2022 seasons \cite[]{mathis2024evaluation}. The COVID-19 pandemic led to the creation of the Health and Human Services (HHS) Patient Impact and Hospital Capacity Data System \cite[]{healthdata2024covidts} which contains COVID-19 and flu hospitalization data, and the COVID-19 Forecast Hub was founded. The COVID-19 Forecast Hub was based on FluSight but with certain major adjustments including how the forecast uncertainty is represented and the addition of the weekly publication of a multi-model ensemble forecast as the official forecast of the CDC \cite[]{bracher2021evaluating, Cramer2022-hub-dataset}. Using estimated quantiles for representing forecast uncertainty and creating a multi-model ensemble are both aspects of the COVID-19 Forecast Hub which were adopted by the flu forecast competition. Additionally the target of the flu forecasts changed from being ILI data to being HHS hospitalization data, which reports the number of hospitalizations due to a laboratory confirmed flu infection  \cite[]{mathis2024evaluation,healthdata2024covidts}. This is as a result of having COVID-19 cases in the population making ILI data, already only a proxy for flu behavior, more difficult to interpret.

@@

The term $\gamma_w$, where $w$ is the season week, is included in (\ref{eq:ili_model}) to capture the per week discrepancy between ILI and the function. Modeling discrepancy has been used in uncertainty analysis of simulators to capture systematic differences between mathematical models and 
reality \cite[]{ma2022multifidelity,brynjarsdottir2014learning,arendt2012improving,kennedy2001bayesian}. Modeling discrepancy can lead to overfitting, particularly in forecasting scenarios, and may also lead to identifiability issues. 
Thus, care must be taken in setting parameter constraints as well as in the selection of prior distributions \cite[]{osthus2019dynamic,brynjarsdottir2014learning}. 
Modeling of the discrepancy for ILI was done by Osthus et al. during the 2015 and 2016 flu seasons where their model outperformed all others in the CDC flu forecasting challenge \cite[]{osthus2019dynamic}. As in Osthus's model, $\gamma_w$ is modeled as a reverse random walk, as shown in (\ref{eq:rw}). 
\begin{equation}
    \label{eq:rw}
        \gamma_w|\gamma_{w + 1} \overset{ind}{\sim} N(\gamma_{t+1},\sigma^2_{\gamma}), \quad
        \gamma_{W} \sim N(0,\sigma^2_{\gamma_W})
\end{equation}
The idea for using the reverse random walk is that there are several previous seasons of ILI data, and assuming the random walk captures systematic behavior, fitting it hierarchically over seasons can assist in predicting future behavior in the current season. 
Reverse random walks have also been used with success in election forecasting and other flu forecasting models \cite[]{osthus2021multiscale, osthus2019dynamic, linzer2013dynamic}. The sum to zero constraint $-\gamma_1 = \sum_{w=2}^W \gamma_w$ is imposed on (\ref{eq:rw}) to improve identifiability.

\subsection{ILI forecasts}
Model (\ref{eq:ili_model}) is fit via Bayesian posterior updating. Future ILI forecasts are obtained via the posterior predictive distribution where for week $w$, the predictive distribution is obtained by integrating over the parameters $\pi_s$, $\kappa_s$, $\sigma^2_{\gamma}$, and $\sigma^2_{\gamma_W}$ as in (\ref{eq:ili_post}) where $p(\boldsymbol{\pi}_s, \kappa_s, \sigma^2_{\gamma}, \sigma^2_{\gamma_W} | \textbf{ILI})$ is the density function of the posterior distribution for the model parameters. If the current week is $w^*$ then the desired forecasts are for weeks $w^* + i$ where $i$ is a positive integer.

\begin{equation}
    \label{eq:ili_post}
    p(\widetilde{ILI}_{s,w^*} | \textbf{ILI}) = \int \int \int \int p(\widetilde{ILI}_{s,w^*} | \boldsymbol{\pi}_s, \kappa_s, \sigma^2_{\gamma}, \sigma^2_{\gamma_W}) p(\boldsymbol{\pi}_s, \kappa_s, \sigma^2_{\gamma}, \sigma^2_{\gamma_W} | \textbf{ILI}) d\boldsymbol{\pi}_s d \kappa_s d \sigma^2_{\gamma} d \sigma^2_{\gamma_W}
\end{equation}
% \jarad{It doesn't make sense to have 4 $\int$ here since $\pi_s$ is a vector.}
% \spencer{Isn't that captured in $d \pi_s$?}

@@

The second component for forecast modeling is based on the relationship between hospitalizations and ILI and is defined in (\ref{eq:final_hosp}).
This is an example of an autoregressive model with exogenous variables where the autoregressive lag is one (ARX(1)) \cite[]{raftery2010online,ljung1987system}. Figure \ref{fig:lag_scatter}
shows scatterplots of the difference between hospitalizations and scaled 1 week 
lags by ILI percentage at the US national level. The points in the left plot are
from hospitalization differences whereas in the right the hospitalizations were
log transformed. The positive and slightly curved relationship suggests model
(\ref{eq:final_hosp}) is a reasonable choice.

\begin{figure}[hbt!]
    \includegraphics[scale=.46]{Images/lagdiff_vs_ili.png}
    \caption{Scatterplots of national hospitalizations (left) or log hospitalizations (right) lag differences by ILI \% where the lag is scaled by $\phi$. Points are colored by season. For hospitalizations $\phi = 0.6$, and for log hospitalizations $\phi = 0.77$. }
    \label{fig:lag_scatter}
\end{figure}

\begin{equation}
    \begin{aligned}
    \label{eq:final_hosp}
    H_{s,w} &= \alpha_{0s} + \alpha_{1s} (ILI_{s,w} \times P) + \alpha_{2s} (ILI_{s,w} \times P)^2 + \phi H_{s,w-1} + \epsilon_{s,w}\\ 
    \epsilon_{s,w} &\overset{iid}{\sim} D_s(0, \sigma_{\epsilon_s}^2 \times P, \omega_s) %N(0, \sigma_{\epsilon}^2)
    \end{aligned}
\end{equation}
In model (\ref{eq:final_hosp}), $H_{s,w}$ is the number of hospitalizations for week $w$ in season $s$, $\epsilon_{s,w}$ is an error term distributed according to some distribution $D_s$ with mean parameter 0, scale parameter $\sigma_{\epsilon_s}$, and the additional parameter $\omega_s$ is the degrees of freedom parameter when $D_s$ belongs to the location-scale t (LST) family.

@@

% \spencer{Blue stuff below will be either included further on or will be removed entirely

@@

The models were fit via Markov chain Monte Carlo (MCMC) sampling using the \texttt{cmdstanr} package which was developed and is maintained by the Stan Development Team \cite[]{stan2024manual} \cite[]{gabry2022stan}. Stan implements Hamiltonian Monte Carlo (HMC) sampling with the No-U-turn sampler \cite[]{hoffman2014no}. The \texttt{cmdstanr} package provides several diagnostic statistics for assessing the sampler.
As mentioned, most model parameter prior distributions were intended to be uninformative.
Plots of posterior distributions for select parameters from ILI and hospitalization models are shown in the supplementary material \cite[]{wadsworth2024bas}.

@@

@@

Model comparison was done by calculating the continuous ranked probability score (CRPS) for each forecast. The CRPS is a proper scoring rule which evaluates the forecast distribution function. Proper scoring rules are the current standard for comparing performance between probabilistic forecasts and selecting the best forecasts according to the notion of maximizing sharpness subject to (auto-)calibration \cite[]{gneiting2007probabilistic, tsyplakov2013evaluation}. Proper scoring rules are commonly used in forecast comparison and have the property that a forecaster is incentivized to be honest in the reporting of their forecasts \cite[]{gneiting2007strictly, gneiting2014probabilistic}. The CRPS is defined in (\ref{eq:crps2}). Here $F(\cdot)$ is the forecast distribution function, and $y^*$ is the observed targeted value of the forecast. The orientation of the CRPS is negative, meaning the smaller the score the better. Here $\vmathbb{1}\{\cdot\}$ is the indicator function.
The CRPS is calculated using a quantile decomposition from \cite[]{laio2007verification}.

\begin{equation}
    \text{CRPS}(F, y^*) = \int_{-\infty}^{\infty} (F(x) - \vmathbb{1}\{y^* \leq x\})^2 dx
    \label{eq:crps2}
\end{equation}

Figure \ref{fig:crps_by_week_horizon} shows 
boxplots of the CRPS for the 500 simulation replicates faceted by week and horizon. 
% Note that in figure \ref{fig:logs_by_week_horizon}, only LogS values of less than 17 are included to improve visualization.
Notable plots here are for week 20 where the two models accounting for the ILI discrepancy, ASGD and SIRD, are the best performing forecast models, with lower medians and lower variation than the two models not accounting for ILI discrepancy. This makes sense given the forecasts at week 20 are forecasting weeks, 22 and 23 where figure \ref{fig:discrepancy} shows a seasonal peak and trough. This confirms that there is value in modeling discrepancy at least during around the holiday weeks near week 22. Among the four models, ASG tends to have the lowest CRPS values though this is not always the case. This suggests that the additional flexibility of the five parameter ASG function is often better for forecasting than that of the three parameter SIR function. Additional
plots in the supplemtary material \cite[]{wadsworth2024bas} further show that the ASG models often but not always outperform the SIR models according to the CRPS and another common proper scoring rule known as the logarithmic score.

\begin{figure}[hbt!]
    \centering
    \includegraphics[scale=.4]{Images/crps_by_week_horizon.png}
    \caption{Boxplots of continuous ranked probability score (CRPS) for the four ILI models over all seasons in the simulation study faceted by horizon (x-axis) and week (y-axis). Horizons include 1-4 week ahead forecasts and weeks include weeks 14, 20, 26, 32, and 38 of the flu season}
    \label{fig:crps_by_week_horizon}
\end{figure}

In this section we apply the forecast models to make forecasts of the 2023 flu season weekly hospitalizations. The scoring of the forecasts is in the context of the FluSight competition where each competing forecast was submitted as a set of 23 quantiles corresponding to the given probability levels $(0.010, 0.025, 0.050, 0.100, 0.150, …, 0.950, 0.975,\\ 0.990)$. A single forecast is thus comprised of 11 predictive intervals and a median.
Forecasts of 1, 2, 3, and 4-week ahead hospitalization counts were requested, and forecasts were made at the state and national level. The first week of forecasting took place during the week of October 7, 2023, and the final week was the week of April 27, 2024 making 29 total weeks of forecasts.
 The same format was used during the 2021 and 2022 seasons and for the COVID-19 Forecast Hub \cite[]{mathis2024evaluation, bracher2021evaluating}. 
% Nearly teams participated and forecasts from over 40 models were submitted throughout the duration of the season. 
 Primary scores for evaluating each forecast were the weighted interval score (WIS), the log-weighted interval score (LWIS), and the relative weighted interval score (RWIS). The focus in this section will be on the LWIS (see supplementary material for RWIS based results \cite[]{wadsworth2024bas}).

The WIS is a proper scoring rule used for scoring quantile or interval forecasts \cite[]{gneiting2007strictly, gneiting2014probabilistic, bracher2021evaluating} and is defined in (\ref{eq:wis2}) where $Q$ is a forecast represented by all included quantiles, $B$ is the number of intervals,  $y^*$ is the observed value targeted by the forecast, $w_0 = 1/2$ and $w_b = \alpha_b / 2$ are weights for each interval, and $\alpha_b$ is the nominal level of the $b^{th}$ interval. $IS_{\alpha}$ is the interval score (IS), a proper scoring rule for a single interval as defined in (\ref{eq:is}). The goal of the forecaster is to minimize the WIS. The LWIS is the same as the WIS except that it is evaluated over the log of quantiles and the log of the observed value.

@@

>>>>

\section{Introduction}

@@

The administration of FluSight saw few changes during the first seven seasons
between 2013 and 2019, 
but the onset of the COVID-19 pandemic and subsequent developments for COVID-19 
forecasting led to major modifications. As a result of the COVID-19 pandemic 
which began during the 2019 flu season, the typical flu outbreak behavior was 
altered between the 2019 and 2022 seasons \cite[]{mathis2024evaluation}. 
The COVID-19 pandemic led to the creation of the Health and Human Services 
(HHS) Patient Impact and Hospital Capacity Data System 
\cite[]{healthdata2024covidts} which contains COVID-19 and flu hospitalization 
data, and the COVID-19 Forecast Hub was founded. The COVID-19 Forecast Hub 
was based on FluSight but with certain major adjustments including how the 
forecast uncertainty is represented and the addition of the weekly publication 
of a multi-model ensemble forecast as the official forecast of the CDC 
\cite[]{bracher2021evaluating, Cramer2022-hub-dataset}. Using estimated 
quantiles for representing forecast uncertainty and creating a multi-model 
ensemble are both aspects of the COVID-19 Forecast Hub which were adopted by 
the flu forecast competition. Additionally the target of the flu forecasts 
changed from being ILI data to being HHS hospitalization data, which reports 
the number of hospitalizations due to a laboratory confirmed flu infection  
\cite[]{mathis2024evaluation,healthdata2024covidts}. This is as a result of 
having COVID-19 cases in the population making ILI data, already only a proxy 
for flu behavior, more difficult to interpret.

@@

The term $\gamma_w$, where $w$ is the season week, is included in 
(\ref{eq:ili_model}) to capture the per week discrepancy between ILI and the 
function. Modeling discrepancy has been used in uncertainty analysis of 
simulators to capture systematic differences between mathematical models and 
reality \cite[]{ma2022multifidelity,brynjarsdottir2014learning,
arendt2012improving,kennedy2001bayesian}. Modeling discrepancy can lead to 
overfitting, particularly in forecasting scenarios, and may also lead to 
identifiability issues. 
Thus, care must be taken in setting parameter constraints as well as in the 
selection of prior distributions \cite[]{osthus2019dynamic,
brynjarsdottir2014learning}. 
Modeling of the discrepancy for ILI was done by Osthus et al. during the 2015 
and 2016 flu seasons where their model outperformed all others in the CDC flu 
forecasting challenge \cite[]{osthus2019dynamic}. As in Osthus's model, 
$\gamma_w$ is modeled as a reverse random walk, as shown in (\ref{eq:rw}). 
\begin{equation}
    \label{eq:rw}
        \gamma_w|\gamma_{w + 1} \overset{ind}{\sim} N(\gamma_{t+1},\sigma^2_{\gamma}), \quad
        \gamma_{W} \sim N(0,\sigma^2_{\gamma_W})
\end{equation}
The idea for using the reverse random walk is that there are several previous 
seasons of ILI data, and assuming the random walk captures systematic behavior, 
fitting it hierarchically over seasons can assist in predicting future behavior 
in the current season. 
Reverse random walks have also been used with success in election forecasting 
and other flu forecasting models \cite[]{osthus2021multiscale, 
osthus2019dynamic, linzer2013dynamic}. The sum to zero constraint 
$-\gamma_1 = \sum_{w=2}^W \gamma_w$ is imposed on (\ref{eq:rw}) to improve 
identifiability.

@@

The second component for forecast modeling is based on the relationship between 
hospitalizations and ILI and is defined in (\ref{eq:final_hosp}).
This is an example of an autoregressive model with exogenous variables where 
the autoregressive lag is one (ARX(1)) 
\cite[]{raftery2010online,ljung1987system}.

\begin{equation}
    \begin{aligned}
    \label{eq:final_hosp}
    H_{s,w} &= \alpha_{0s} + \alpha_{1s} (ILI_{s,w} \times P) + \alpha_{2s} 
    (ILI_{s,w} \times P)^2 + \phi H_{s,w-1} + \epsilon_{s,w}\\ 
    \epsilon_{s,w} &\overset{iid}{\sim} D_s(0, \sigma_{\epsilon_s}^2 \times P, 
    \omega_s) %N(0, \sigma_{\epsilon}^2)
    \end{aligned}
\end{equation}

Figure \ref{fig:lag_scatter} helps
illustrate why this modeling decision was made.
shows scatterplots of ILI percentage and hospitalizations on the top row.
The bottom row shows scatterplots of the difference between hospitalizations and 
scaled 1 week 
lags by ILI percentage at the US national level. The plots in the right column
are from the logarithm of hospitalizations. The lines through the points is
drawn by connecting hospitalizations for consecutive weeks during the flu 
season.
These plots show that a linear or quadratic relationship between ILI percentage 
and hospitalizations does not capture much of the temporal correlation in 
hospitalizations as shown by the obvious loops drawn in the plot. 
However, when including a 
single lag of hospitalizations in the model, much of the time correlation is
accounted for as shown by the tighter and more frequent crossing of lines shown
in the bottom row. Higher than order 1 autoregressions showed little or no 
improvement over order 1, hence we consider the ARX(1) reasonably well 
suited for modeling ILI and hospitalzations.

\begin{figure}[hbt!]
    \includegraphics[scale=.2]{Images/ili_v_svat_vs_lagdiff.png}
    \caption{Scatterplots of national hospitalizations (left) or log 
    hospitalizations (right) lag differences by ILI \% where the lag is 
    scaled by $\phi$. Points are colored by season. For hospitalizations 
    $\phi = 0.6$, and for log hospitalizations $\phi = 0.77$. }
    \label{fig:lag_scatter}
\end{figure}

In model (\ref{eq:final_hosp}), $H_{s,w}$ is the number of hospitalizations for 
week $w$ in season $s$, $\epsilon_{s,w}$ is an error term distributed 
according to some distribution $D_s$ with mean parameter 0, scale parameter 
$\sigma_{\epsilon_s}$, and the additional parameter $\omega_s$ is the degrees 
of freedom parameter when $D_s$ belongs to the location-scale t (LST) family.

@@

\subsection{Hospitalization Forecasts Using the Bayesian Cut}

Rather than combining the ILI model in equation (\ref{eq:ili_model}) with the 
hospitalization model in equation (\ref{eq:final_hosp}) and computing a fully Bayesian
posterior distribution, both models were fit separately and 1 to 4 week ahead
ILI forecasts from model (\ref{eq:ili_model}) were combined with the posterior
predictive distribution of model (\ref{eq:final_hosp}) to obtain 1 to 4 week
ahead hospitalization forecasts. In this scheme, the posterior distribution of 
model (\ref{eq:ili_model}) parameters was not directly influenced by the 
hospitalization data and the posterior distribution of model 
(\ref{eq:final_hosp}) parameters was not directly influenced by model 
(\ref{eq:ili_model}) parameters. Hence there was limited feedback between the
two components, but propagation of uncertainty in ILI forecasts was still 
permitted in the estimating of hospitalization forecasts.

This two-component modeling is remeniscent of the "Bayesian cut" as 
described in \cite{plummer2015cuts} and \cite{nott2023bayesian}. Cutting a model
into two component models or "modules" may be justfied for several reasons
including to reduce time to fit the model, avoiding mixing issues of posterior
sampling distributions, improving predictive performance or to prevent
model misspecification in one model to influence the other 
\cite[]{nott2023bayesian, jacob2020unbiased, jacob2017better, plummer2015cuts}.

The decision to cut the model into two components was made for several reasons.
When fitting the fully Bayesian joint model, the time to fit a single model using
Markov chain Monte Carlo (MCMC) sampling and
obtain forecasts was sometimes upwards of 4 hours whereas fitting the two models
separately usually took less than 1 hour. When assessing convergence of model
parameters, it was rare that posterior sampling chains for the joint model
showed convergence even after 55,000 iterations, whereas the component model 
fits gave no signs to be concerened with convergence. And most importantly,
the forecasts from the component models generally outperformed those of the 
joing model in terms of minimizing proper scoring rules. 
The process for obtaining forecasts is described in the following paragraphs 
and computation is described in section \ref{sec:implementation_posterior}.

Model (\ref{eq:ili_model}) is fit via Bayesian posterior updating. Future ILI 
forecasts are obtained via the posterior predictive distribution where for
week $w$, the predictive distribution is obtained by integrating over the 
parameters $\pi_s$, $\kappa_s$, $\sigma^2_{\gamma}$, and $\sigma^2_{\gamma_W}$ 
as in (\ref{eq:ili_post}) where 
$p(\boldsymbol{\pi}_s, \kappa_s, \sigma^2_{\gamma}, \sigma^2_{\gamma_W} | \textbf{ILI})$ 
is the density function of the posterior distribution for the model parameters. 
If the current week is $w$ then the desired forecasts are for weeks $w + i$ 
where $i$ is a positive integer. Here $\widetilde{ILI}_{s,w + i}$ represents the 
predicted or forecast ILI at $i$ weeks in the future.

\begin{equation}
    \label{eq:ili_post}
    p(\widetilde{ILI}_{s,w +i} | \textbf{ILI}) = \int \int \int \int 
    p(\widetilde{ILI}_{s,w + i} | \boldsymbol{\pi}_s, \kappa_s, 
    \sigma^2_{\gamma}, \sigma^2_{\gamma_W}) p(\boldsymbol{\pi}_s, \kappa_s, 
    \sigma^2_{\gamma}, \sigma^2_{\gamma_W} | \textbf{ILI}) 
    d\boldsymbol{\pi}_s d \kappa_s d \sigma^2_{\gamma} d \sigma^2_{\gamma_W}
\end{equation}
% \jarad{It doesn't make sense to have 4 $\int$ here since $\pi_s$ is a vector.}
% \spencer{Isn't that captured in $d \pi_s$?}
Hospitalization forecasts $\widetilde{H}_{s,w + i}$ at week $w + i$ are then
obtained by combining the posterior predictive distribution of model 
(\ref{eq:final_hosp}) with (\ref{eq:ili_post}) and integrating over 
$\widetilde{ILI}_{s,w + i}$ as in (\ref{eq:hosp_post_forc}). Here 
$\boldsymbol{\alpha}$ represents all regression and variance parameters in
model (\ref{eq:final_hosp}).

\begin{equation}
\begin{aligned}
\label{eq:hosp_post_forc}
    p(\tilde{H}_{s,w + i} | \textbf{ILI}, \textbf{H}, \boldsymbol{\alpha}) &= 
       \int p(\tilde{H}_{s,w + i} | \textbf{H}, \textbf{ILI}, 
       \widetilde{ILI}_{s,w + i}, \boldsymbol{\alpha}) 
       p(\widetilde{ILI}_{s,w + i} | \textbf{ILI}) d\widetilde{ILI}_{s,w + i}
\end{aligned}
\end{equation}

@@

The models were fit via MCMC sampling using the 
\texttt{cmdstanr} package which was developed and is maintained by the Stan 
Development Team \cite[]{stan2024manual} \cite[]{gabry2022stan}. Stan 
implements Hamiltonian Monte Carlo (HMC) sampling with the No-U-turn 
sampler \cite[]{hoffman2014no}. The \texttt{cmdstanr} package provides 
several diagnostic statistics for assessing the sampler.
As mentioned, most model parameter prior distributions were intended to be 
uninformative.
Plots of posterior distributions for select parameters from ILI and 
hospitalization models are shown in the supplementary material 
\cite[]{wadsworth2024bas}.

@@

Besides the 4 ILI models, we also included forecasts for a naive baseline
(BASE) model and an autoregressive integrated moving average (ARIMA) model. 
The BASE model is a simple random walk with drift model, and the ARIMA model
is fit by selecting the best possible model give the data according to a 
fitting criterion. Both the BASE and ARIMA ILI models were fit using the 
\texttt{forecast} package in \texttt{R} \cite{hyndman2008forecast}. Samples
were then drawn from the forecast distributions, and these were then used
along with the model in (\ref{eq:final_hosp}) to make hospitalization 
forecast distributions.

@@

Model comparison was done by calculating a proper scoring rule for each 
forecast.
Proper scoring rules are the 
current standard for comparing performance between probabilistic forecasts 
and selecting the best forecasts according to the notion of maximizing 
sharpness subject to (auto-)calibration 
\cite[]{gneiting2007probabilistic, tsyplakov2013evaluation}. Proper scoring 
rules are commonly used in forecast comparison and have the property that a 
forecaster is incentivized to be honest in the reporting of their forecasts 
\cite[]{gneiting2007strictly, gneiting2014probabilistic}.
The proper scoring rule primarily used in the FluSight competition and the 
COVID-19 Forecast Hub and which we will use herein is the 
weighted interval score (WIS)
\cite[]{mathis2024evaluation, bracher2021evaluating}.
The WIS is used for scoring quantile or interval 
forecasts 
\cite[]{gneiting2007strictly, gneiting2014probabilistic, bracher2021evaluating} 
and is defined in (\ref{eq:wis2}) where $Q$ is a forecast represented by all 
included quantiles, $B$ is the number of intervals,  $y^*$ is the observed 
value targeted by the forecast, $w_0 = 1/2$ and $w_b = \alpha_b / 2$ are 
weights for each interval, and $\alpha_b$ is the nominal level of the $b^{th}$ 
interval. $IS_{\alpha}$ is the interval score (IS), a proper scoring rule for a 
single interval as defined in (\ref{eq:is}). Here 
$\vmathbb{1}\{\cdot\}$ is the indicator function.
The goal of the forecaster is to 
minimize the WIS.

@@

%  The CRPS is defined 
% in (\ref{eq:crps2}). Here $F(\cdot)$ is the forecast distribution function, 
% and $y^*$ is the observed targeted value of the forecast. The orientation of 
% the CRPS is negative, meaning the smaller the score the better. Here 
% $\vmathbb{1}\{\cdot\}$ is the indicator function.
% The CRPS is calculated using a quantile decomposition from 
% \cite[]{laio2007verification}. 
% 
% \begin{equation}
%     \text{CRPS}(F, y^*) = \int_{-\infty}^{\infty} (F(x) - \vmathbb{1}\{y^* \leq x\})^2 dx
%     \label{eq:crps2}
% \end{equation}

Figure \ref{fig:wis_by_week_horizon} shows 
boxplots of the WIS for the 500 simulation replicates faceted by season week 
on the y-axis and horizon on the x-axis. 
% Note that in figure \ref{fig:logs_by_week_horizon}, only LogS values of less than 17 are included to improve visualization.
In most plots the ASGD model shows the best performance, 
especially during week 20 where
forecasts are near weeks 22 and 23 which show a seasonal peak and trough in 
figure \ref{fig:discrepancy}. The SIRD also shows improved performance over the
SIR model, though its performance does not match that of the ASG or ASGD.
Here the ASGD model regularly outperforms
the BASE and ARIMA models, but the three other nonlinear models do not. 
The ASG model
tends to outperform the SIR and SIRD models suggesting that the additional 
flexibility of the five parameters ASG function 
is often better for forecasting than that of the three parameter SIR function.
But the best performance comes with the combination of the ASG function and
the discrepancy modeling.

% Notable plots here are for week 20 where the two models accounting for the 
% ILI discrepancy, ASGD and SIRD, are the best performing forecast models, with 
% lower medians and lower variation than the two models not accounting for ILI 
% discrepancy. This makes sense given the forecasts at week 20 are forecasting 
% weeks, 22 and 23 where figure \ref{fig:discrepancy} shows a seasonal peak and 
% trough. This confirms that there is value in modeling discrepancy at least 
% during around the holiday weeks near week 22. Among the four models, ASG 
% tends to have the lowest CRPS values though this is not always the case. 
% This suggests that the additional flexibility of the five parameter ASG 
% function is often better for forecasting than that of the three parameter SIR 
% function. 
% Additional
% plots in the supplemtary material \cite[]{wadsworth2024bas} further show that 
% the ASG models often but not always outperform the SIR models according to the 
% CRPS and another common proper scoring rule known as the logarithmic score.
